### 2021-06-04 analysis/index.Rmd
chore: initial wflow site commit

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Please find all project details in the [About](about.html) page.