### 2021-06-11 analysis/license.Rmd
feat: add basic targets workflow

---
title: "License"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

**Code** is released under the [GNU GPLv3 license][gpl].

**Text** is released under the
[Creative Commons Attribution 4.0 International License][cc].

[gpl]: https://choosealicense.com/licenses/gpl-3.0/
[cc]: https://creativecommons.org/licenses/by/4.0/