### 2021-06-11 analysis/missing.Rmd
feat: plot missing values

---
title: "Missing Data"
output:
  workflowr::wflow_html:
    toc: true
    fig.width: 10
    fig.height: 10
editor_options:
  chunk_output_type: console
---

```{r missing_setup}
library("targets")
tar_load(raw_data)

.dotplot <- function(x, xlim = c(0, length(x)),
                     col = palette.colors(2L)[2L], pch = 19L) {
    old.par <- par(no.readonly = TRUE)
    on.exit(par(old.par))

    n <- length(x)

    mai <- par("mai")
    w <- max(strwidth(names(x), "inch"), na.rm = TRUE) + 1/16
    if (mai[2L] < w)
        mai[2L] <- mai[4L] + w # taken from dotchart
    par(mai = mai)

    plot.window(xlim = xlim, ylim = c(0L, n + 1L))
    title(main = "Missing Data", adj = 0L)
    title(xlab = "Frequency", adj = 1L)
    y <- seq_len(n)
    mtext(
        names(x), at = y, adj = 0L, side = 2L, las = 2L,
        line = (w + 0.1) / par("csi"), cex = 0.8
    )
    abline(h = y, col = "#808080", lty = "dotted", lwd = 1L)
    points(x, y, col = col, pch = pch)
    axis(1L, lwd.ticks = 0L, col = "#808080")
}

.image <- function(x, col = c("white", palette.colors(2L)[2L])) {
    old.par <- par(no.readonly = TRUE)
    on.exit(par(old.par))

    nr <- nrow(x)
    nc <- ncol(x)
    nms <- colnames(x)

    mai <- par("mai")
    w <- max(strwidth(nms, "inch"), na.rm = TRUE) + 1/16
    if (mai[2L] < w)
        mai[2L] <- mai[4L] + w # taken from dotchart
    par(mai = mai)

    y <- seq_len(nc)
    image(
        1L:nr, y, x[, rev(y)], col = col,
        xlim = 0.5 + c(0L, nr), ylim = 0.5 + c(0L, nc),
        axes = FALSE, xlab = "", ylab = ""
    )
    title(main = "Missing Data", adj = 0L)
    title(xlab = "Rows", adj = 1L)
    mtext(
        rev(nms), at = y, adj = 0L, side = 2L, las = 2L,
        line = (w + 0.1) / par("csi"), cex = 0.8
    )
    abline(h = y, col = "#808080", lty = "dotted", lwd = 1L)
}
```

# Missing Data

```{r missing_table}
m <- raw_data
m$Sex <- as.numeric(m$Sex)
m <- as.matrix(m)
mna <- is.na(m)
mode(mna) <- "numeric"
tbl <- colSums(mna)
na <- sort(tbl[tbl > 0L], decreasing = TRUE)
knitr::kable(
    cbind(na, round(colMeans(mna[, names(na)]) * 100, 2)),
    col.names = c("# Missing Data", "% of Missing Data")
)

.dotplot(sort(na), xlim = c(0L, nrow(m)))
```

```{r missing_image, fig.height = 10, fig.width = 12}
.image(mna)
```