### 2021-06-11 analysis/histograms.Rmd

---
title: "Histograms"
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r histograms_setup, message = FALSE}
library("targets")
tar_load(imp_data)
tar_load(zlog_data)

.hist <- function(
    ..., xlab = deparse(substitute(x)), main = paste("Histogram of", xlab),
    col = palette.colors(2L)[2L]
)hist(..., main = main, xlab = xlab, col = col)

.qqnorm <- function(y, ..., yname = deparse(substitute(y)),
    main = paste0("Normal Q-Q Plot (", yname, ")"),
    ylab = paste0("Sample Quantiles (", yname, ")"),
    col = palette.colors(2L)[2L]) {
    qqnorm(y, ..., main = main, ylab = ylab, col = col, pch = 19L)
    qqline(y, ..., col = col)
}
```

# Age

```{r histograms_age}
.hist(imp_data$Age, xlab = "Age")
```

# Raw Data (imputed)

```{r histograms_labs, results = "asis", message = FALSE, warning = FALSE}
labv <- sort(grep("_[SECQ]$", colnames(imp_data), value = TRUE))

for (v in labv) {
    cat("\n\n##", v , "\n\n")
    .hist(imp_data[[v]], main = paste0("Histogram of ", v), xlab = v)
    .qqnorm(imp_data[[v]], yname = v)
}
```

# Zlog Data (imputed)

```{r histograms_labs_zlog, results = "asis", message = FALSE, warning = FALSE}
for (v in labv) {
    cat("\n\n##", v , "\n\n")
    .hist(zlog_data[[v]], main = paste0("Histogram of ", v), xlab = v)
    .qqnorm(zlog_data[[v]], yname = v)
}
```