### 2021-07-02 analysis/index.Rmd
refactor: remove about page

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

# Acknowledgment

This work is part of the [AMPEL](https://ampel.care/en/)
(Analysis and Reporting System for the Improvement of Patient Safety through
Real-Time Integration of Laboratory Findings) project.

This measure is co-funded with tax revenues based on the budget adopted by
the members of the Saxon State Parliament.