### 2021-07-02 analysis/index.Rmd
feat: add first citations

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
bibliography:
  - "`r rprojroot::find_rstudio_root_file('analysis', 'bibliography', 'bibliography.bib')`"
  - "`r rprojroot::find_rstudio_root_file('analysis', 'bibliography', 'rpackages', 'home.bib')`"
---

```{r, include=FALSE}
knitr::write_bib(
    c("targets", "workflowr"),
    file = rprojroot::find_rstudio_root_file(
        "analysis", "bibliography", "rpackages", "home.bib"
    )
)
```

## Acknowledgment

### AMPEL

This work is part of the [AMPEL](https://ampel.care/en/)
(Analysis and Reporting System for the Improvement of Patient Safety through
Real-Time Integration of Laboratory Findings) project [@eckelt2020].

This measure is co-funded with tax revenues based on the budget adopted by
the members of the Saxon State Parliament.

### R Packages

This site is generated by combining the
[workflowr](https://jdblischak.github.io/workflowr/) [@workflowr2019] and the
[targets](https://docs.ropensci.org/targets/) [@targets2021] package.

A visual network of the pipeline is depicted [here](pipeline.html).

## References