### 2021-08-02 analysis/license.Rmd

---
title: "License"
output:
  bookdown::html_document2:
    toc: false
---

**Code** is released under the [GNU GPLv3 license][gpl].

**Text** is released under the
[Creative Commons Attribution 4.0 International License][cc].

[gpl]: https://choosealicense.com/licenses/gpl-3.0/
[cc]: https://creativecommons.org/licenses/by/4.0/