### 2021-08-02 analysis/benchmarks.Rmd
fix: working directory for targets

---
title: "Benchmarks"
output:
  bookdown::html_document2:
    fig.width: 12
    fig.height: 8
---

```{r benchmarks_setup, message = FALSE}
library("targets")
library("mlr3viz")
library("viridis")
library("ggplot2")

withr::with_dir(rprojroot::find_rstudio_root_file(), {
    tar_load(bmrk_results)
})
```

```{r benchmarks_plot_c_harrell}
autoplot(bmrk_results) +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```