### 2021-08-02 analysis/corrplot.Rmd

---
title: "Corrplot"
output:
  bookdown::html_document2:
    fig.width: 12
    fig.height: 12
---

```{r corrplot_setup, message = FALSE}
library("corrplot")
library("targets")

withr::with_dir(rprojroot::find_rstudio_root_file(), {
    tar_load(imp_data)
})
```

# Correlation

```{r corrplot}
labv <- grep("_[SECQ]$", colnames(imp_data), value = TRUE)
cr <- cor(imp_data[labv], use = "pairwise.complete.obs")
corrplot.mixed(
    cr,
    tl.pos = "lt", tl.col = "black", tl.cex = 0.75,
    number.cex = 0.25, number.font = 1L,
    order = "AOE"
)
```