### 2021-08-02 analysis/tables.Rmd

---
title: "Tables"
---

```{r histograms_setup, message = FALSE}
library("targets")
library("skimr")

withr::with_dir(rprojroot::find_rstudio_root_file(), {
    tar_load(raw_data)
    tar_load(imp_data)
    tar_load(zlog_data)
})
```

# Raw Data

```{r tables_skim_raw}
skim(raw_data)
```

# Raw Data (imputed)

```{r tables_skim_imp}
skim(imp_data)
```

# Zlog Data

```{r tables_skim_zlog}
skim(zlog_data)
```