### 2021-08-04 analysis/tables.Rmd
feat: add table 1 and table observed vs expected mortality

---
title: "Tables"
---

```{r histograms_setup, message = FALSE}
library("targets")
library("skimr")
library("zlog")
tar_load(raw_data)
tar_load(imp_data)
tar_load(zlog_data)
```

# Raw Data

```{r tables_skim_raw}
skim(raw_data)
```

# Raw Data (imputed)

```{r tables_skim_imp}
skim(imp_data)
```

# Zlog Data

```{r tables_skim_zlog}
skim(zlog_data)
```

# Raw Data (imputed) with Zlog Colors

```{r tables_raw_zlog, results = "asis", echo = FALSE, message = FALSE}
k <- kableExtra::kable_classic(
    kableExtra::kbl(
        raw_data,
        digits = 2,
        escape = FALSE
    ),
    "basic"
)
for (i in grep("_[CEQS]$", colnames(raw_data)))
    k <- kableExtra::column_spec(k, i, background = zcol(zlog_data[[i]]))
k |> kableExtra::scroll_box(width = "100%", height = "500px")
```

# Table One

```{r tbl1, results = "asis", echo = FALSE, message = FALSE}
tar_read(tbl1)
```