### 2021-10-15 analysis/benchmarks.Rmd
refactor: reset figure width

---
title: "Benchmarks"
output:
  bookdown::html_document2:
    fig.width: 18
    fig.height: 8
---

```{r benchmarks_setup, message = FALSE}
library("targets")
library("mlr3viz")
library("viridis")
library("ggplot2")
tar_load(bmrk_results)
tar_load(bmrk_aggr)

p.value <- 0.2
```

# Overview

```{r benchmarks_plot_c_harrell, fig.width = 18}
autoplot(bmrk_results) +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

# Aggregated performance

Aggregated performance across all `nrow(bmrk_results$tasks)` datasets.

```{r benchmarks_aggr_plot_c_harrell}
autoplot(bmrk_aggr, type = "box", meas = "harrell_c") +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

# Pairwise comparison

*p*-value is set to `r p.value`
(otherwise an error is thrown because of insignificant results).

```{r benchmarks_aggr_pairwise_harrell_c}
autoplot(
    bmrk_aggr, type = "fn", meas = "harrell_c", p.value = p.value,
    col = palette.colors(2L)[2L]
)
```

# Critical difference

*p*-value is set to `r p.value`
(otherwise an error is thrown because of insignificant results).

```{r benchmarks_aggr_critdiff_c_harrell}
autoplot(
    bmrk_aggr, type = "cd", meas = "harrell_c", style = 2, p.value = p.value,
    minimize = FALSE
) + scale_color_viridis(discrete = TRUE)
```