### 2022-03-17 analysis/elasticnet.Rmd
feat: add elastic net bootstrap and timeROC evaluations

---
title: "elasticnet"
output:
  bookdown::html_document2:
    fig.width: 18
    fig.height: 8
---

```{r elasticnet_setup, message = FALSE}
library("targets")
library("ameld")
library("viridisLite")
tar_load(arcvob)
tar_load(bootrcv)
tar_load(bootarcv)
#tar_load(bootarcv9)
tar_load(amelddata)
tar_load(ameldcfg)
tar_load(zlog_data)
```

# Tuning alpha

```{r arcvob}
arcvob
```

```{r arcvob-plot-alpha}
plot(arcvob)
```

```{r arcvob-plot-lambdamin}
plot(arcvob, what = "lambda.min")
```

```{r arcvob-plot-lambda1se}
plot(arcvob, what = "lambda.1se")
```

# Bootstrapping

## rcv.glmnet

```{r bootrcv}
plot(bootrcv, what = "calibration")

ps <- lapply(
    zlog_data[paste0("SurvProbMeld", c("Unos", "NaUnos", "Plus7"))],
    function(p) {
        ctpnts <- cutpoints(p, n = ameldcfg$m)
        f <- cut(p, ctpnts, include.lowest = TRUE)
        list(
            predicted = groupmean(p, f = f),
            observed = observed_survival(
                amelddata$y, f = f, times = ameldcfg$times
            )
        )
    }
)
names(ps) <- c("MELD", "MELD-Na", "MELD-Plus7")
col <- viridisLite::viridis(6)[4:6]

for (i in seq_along(ps)) {
    lines(
        ps[[i]]$predicted, ps[[i]]$observed, col = col[i], type = "b", pch = 19
    )
}
legend("topleft", col = col, legend = names(ps), pch = 19, bty = "n")

plot(bootrcv, what = "selected", cex = 0.5)
plot(bootrcv$fit$glmnet.fit, xvar = "norm")
plot(bootrcv$fit$glmnet.fit, xvar = "lambda")
plot(bootrcv$fit$glmnet.fit, xvar = "dev")
```

## arcv.glmnet

```{r bootarcv}
a <- c(table(sapply(bootarcv$models, function(m)m$fit$alpha)))
plot(bootarcv, what = "calibration")
```
```{r bootarcv2, fig.height = 12}
plot(bootarcv, what = "selected")
plot_dots(a, main = "Selected Alpha Values")
```

## arcv9.glmnet

```{r bootarcv9}
#a <- c(table(sapply(bootarcv9$models, function(m)m$fit$alpha)))
#plot(bootarcv9 , what = "calibration")
#plot(bootarcv9, what = "selected")
#plot_dots(a, main = "Selected Alpha Values")
```