### 2022-03-17 analysis/corrplot.Rmd
feat: list correlations in table

---
title: "Corrplot"
output:
  bookdown::html_document2:
    fig.width: 12
    fig.height: 12
---

```{r corrplot_setup, message = FALSE}
library("corrplot")
library("targets")
tar_load(imp_data)
```

# Correlation

## Plot

```{r corrplot}
labv <- grep("_[SECQ]$", colnames(imp_data), value = TRUE)
cr <- cor(imp_data[labv], use = "pairwise.complete.obs")
corrplot.mixed(
    cr,
    tl.pos = "lt", tl.col = "black", tl.cex = 0.75,
    number.cex = 0.25, number.font = 1L,
    order = "AOE"
)
```

## Table

```{r cortbl}
ut <- upper.tri(cr)
nms <- paste(
    colnames(cr)[col(cr)[ut]],
    rownames(cr)[row(cr)[ut]],
    sep = ":"
)
tbl <- data.frame(Values = nms, Correlation = cr[ut])
tbl <- tbl[order(-abs(tbl$Correlation)),]
knitr::kable(tbl, header = TRUE, row.names = FALSE, digits = 3L)
```