### 2022-06-06 analysis/benchmarks.Rmd
fix: bootstraping elastic net

---
title: "Benchmarks"
output:
  bookdown::html_document2:
    fig.width: 18
    fig.height: 8
---

```{r benchmarks_setup, message = FALSE}
library("targets")
library("data.table")
library("mlr3")
library("mlr3misc")
library("mlr3proba")
library("mlr3viz")
library("viridis")
library("ggplot2")
tar_load(bmrk_results)
tar_load(bmrk_aggr)

agg <- bmrk_results$aggregate(msr("surv.cindex", id = "harrell"))
agg[, `:=`
    (nr = NULL, resample_result = NULL, resampling_id = NULL, iters = NULL)
]
setorder(agg, task_id, -harrell)
p.value <- 0.05
```

# Overview

## Table

```{r benchmarks_table, results = "asis"}
lapply(split(agg, agg$task_id), knitr::kable, digits = 4)
```

## Boxplots

```{r benchmarks_plot_c_harrell, fig.width = 18, fig.height = 12}
m <- agg[, max(harrell), by = task_id]
autoplot(bmrk_results) +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE) +
    geom_hline(
        aes(yintercept = V1), linetype = "dashed", color = "red", data = m
    )
```

```{r benchmarks_plot_c_uno, fig.width = 18, fig.height = 12}
autoplot(
    bmrk_results,
    measure = msr("surv.cindex", id = "uno", weight_meth = "G2")) +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE)
```

# Aggregated performance

Aggregated performance across all `r nrow(bmrk_results$tasks)` datasets.

```{r benchmarks_aggr_plot_c_harrell}
autoplot(bmrk_aggr, type = "box", meas = "harrell") +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE)
```

```{r benchmarks_aggr_plot_c_uno}
autoplot(bmrk_aggr, type = "box", meas = "uno") +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE)
```

# Pairwise comparison

*p*-value is set to `r p.value`
(otherwise an error is thrown because of insignificant results).

```{r benchmarks_aggr_pairwise_harrell_c}
autoplot(
    bmrk_aggr, type = "fn", meas = "harrell", p.value = p.value,
    col = palette.colors(2L)[2L]
)
```

# Critical difference

*p*-value is set to `r p.value`
(otherwise an error is thrown because of insignificant results).

```{r benchmarks_aggr_critdiff_c_harrell}
autoplot(
    bmrk_aggr, type = "cd", meas = "harrell", style = 2, p.value = p.value,
    minimize = FALSE
) + scale_color_viridis(discrete = TRUE)
```