### 2022-07-17 analysis/article.Rmd
refactor: integrate Thorsten's review comments

---
# R markdown template taken from
# Dan Holmes <dtholmes@mail.ubc.ca>
# https://labrtorian.com/2019/08/26/rmarkdown-template-that-manages-academic-affiliations/
title: "A New Model to Predict Survival in Patients with End-Stage Liver Disease"
subtitle: >
    Comparision against MELD, MELD-Na, MELD-Plus7.
abstract: >
  **Objectives**:
      The shortage in grafts for liver transplantation requires
      risk stratification and adequate allocation rules.
      This study aims to improve the model of end-stage liver disease (MELD)
      score for 90-day mortality prediction with the help of different
      machine learning (ML) algorithms.

  **Methods**:
      We retrospectively analyzed clinical and laboratory data of 654
      patients who were recruited during the evaluation process for
      liver transplantation at the University Hospital Leipzig.
      After comparing 13 different ML algorithms in a nested cross-validation
      setting and selecting the best performing one we build a new model
      to predict 90-day mortality in patients with end-stage liver disease.

  **Results**:
      Penalized regression algorithms yield the highest prediction performance
      in our ML algorithm benchmark.
      In favour of a simpler model we chose the least absolute shrinkage and
      selection operator (lasso) regression.
      Beside the classical MELD variables INR and bilirubin the
      lasso regression selected cystatin C over creatinine, and
      additionally IL-6, total protein and cholinesterase.
      The new model offers an improved discrimination and calibration than MELD
      and MELD with sodium (MELD-Na) or the MELD-Plus7 risk score.

  **Conclusion**:
      We provide a new ML based model of end-stage liver disease that
      outperforms the classical MELD score for 90-day survival prediction.

  **Keywords**:
      end-stage liver disease; MELD; risk stratification; machine learning;
      survival estimation; IL-6.

author:
    - Sebastian Gibb:
        email: mail@sebastiangibb.de
        institute: [ILM, KAINS]
        correspondence: true
    - Thorsten Kaiser:
        email: thorsten.kaiser@medizin.uni-leipzig.de
        institute: ILM
        correspondence: false
institute:
    - ILM: Institute of Laboratory Medicine, Clinical Chemistry and Molecular Diagnostics, University Hospital Leipzig, Paul-List-Str. 13-15, D-04103 Leipzig, Germany.
    - KAINS: Anesthesiology and Intensive Care Medicine, University Hospital Greifswald, Ferdinand-Sauerbruch-Straße, D-17475 Greifswald, Germany.
date: "`r Sys.Date()`"
bibliography:
  - "`r rprojroot::find_rstudio_root_file('analysis', 'bibliography', 'bibliography.bib')`"
  - "`r rprojroot::find_rstudio_root_file('analysis', 'bibliography', 'rpackages', 'article.bib')`"
  - "`r rprojroot::find_rstudio_root_file('analysis', 'bibliography', 'rpackages', 'ameld.bib')`"
  - "`r rprojroot::find_rstudio_root_file('analysis', 'bibliography', 'rpackages', 'zlog.bib')`"
output:
  bookdown::html_document2:
    pandoc_args: # pandoc_args doesn't support r evaluation
      - --lua-filter=pandoc/lua-filters/scholarly-metadata.lua
      - --lua-filter=pandoc/lua-filters/author-info-blocks.lua
  bookdown::word_document2:
    pandoc_args: # pandoc_args doesn't support r evaluation
      - --lua-filter=pandoc/lua-filters/scholarly-metadata.lua
      - --lua-filter=pandoc/lua-filters/author-info-blocks.lua
---

```{r knitr_setup, include=FALSE}
knitr::write_bib(
    c(
        "base",
        "targets", "workflowr",
        "gtsummary",
        "mlr3", "mlr3proba",
        "survival", "glmnet", "penalized", "ranger", "randomForestSRC",
        "xgboost", "survivalsvm", "survivalmodels", "timeROC"
    ),
    file = rprojroot::find_rstudio_root_file(
        "analysis", "bibliography", "rpackages", "article.bib"
    )
)
cat(
    gsub("@Manual\\{", "@Manual\\{R-ameld", toBibtex(citation("ameld"))),
    file = rprojroot::find_rstudio_root_file(
        "analysis", "bibliography", "rpackages", "ameld.bib"
    )
)
cat(
    gsub("@Manual\\{", "@Manual\\{R-zlog", toBibtex(citation("zlog"))),
    file = rprojroot::find_rstudio_root_file(
        "analysis", "bibliography", "rpackages", "zlog.bib"
    )
)
```

```{r article_setup, include=FALSE}
library("targets")
library("ggplot2")
library("mlr3viz")
library("viridis")
library("ameld")
library("timeROC")

tar_load(bmrk_results)
tar_load(crossval)
tar_load(tuner)

tar_load(raw_data)
tar_load(labelled_meld_data)
tar_load(zlog_data)

tar_load(tbl1)
tar_load(tbl_observed_vs_expected_mortality)

tar_load(ameldcfg)
tar_load(amelddata)
tar_load(rngr)

tar_load(timeROC_MELD)
tar_load(timeROC_MELDNa)
tar_load(timeROC_MELDPlus7)
tar_load(timeROC_RCV)
```

# Introduction

Liver cirrhosis is the terminal result of fibrotic remodeling
of liver tissue due to chronic damage.
This end-stage of liver disease is usually irreversible and
the only available therapy is liver transplantation.
However, the shortage in grafts for transplantation from deceased donors
requires risk stratification and precise and fair allocation rules.
The allocation of liver transplantation in most countries is based on
disease severity determined by the model of end-stage liver disease (MELD)
[@malinchoc2000; @wiesner2003; @optn_policies2021].
The MELD score estimates the patients' 3-month mortality risk based on
laboratory results, namely bilirubin, creatinine and international normalized
ratio (INR).
The MELD score has been extended by the sodium level
(MELD-Na score) as this was found to be an important additional risk factor in
liver cirrhosis [@kim2008; @optn_policies2021].
The MELD was initially developed for predicting the survival of patients
undergoing transjugular intrahepatic portosystemic shunts.
Afterwards it was revalidated for predicting mortality risk in patients awaiting
a liver transplantation.
Although the MELD score should be an objective allocation score especially the
creatinine and INR are highly dependent on the utilized laboratory methods
[@trotter2004; @cholongitas2007]. Patients with identical disease state could
have very different MELD scores and thus get different priority on the liver
transplantation waiting list.
Furthermore often, e.g. for acute-on-chronic liver failure, the MELD score
underestimates the mortality risk [@hernaez2020].

There have been some attempts to use the data extracted from more than 300.000
electronic medical records from two hospitals in the United States to improve
the MELD score [@kartoun2017]. The derived MELD-Plus7 and MELD-Plus9 risk score
add albumin, white blood cell count, total cholesterol, age and length of
stay to the MELD-Na variables. Despite its published prediction improvement the
MELD-Plus scores are not used for transplant allocation yet.

As depicted by MELD-Plus risk scores better predictive scores often need more
variables and are more complicated. To reduce the risk of overlooking or
incorrectly calculating and interpreting the results,
clinical decision support systems may be used and could improve patient safety.
The research project on digital laboratory medicine (AMPEL) develops a
clinical decision support system based on laboratory diagnostics that should
support clinical practitioners in interpreting the laboratory results and taking
the necessary medical interventions [@eckelt2020].

This study aims to find clinical and laboratory values that improve the risk
stratification for liver transplantation over classical MELD, MELD-Na and
MELD-Plus scores and could be implemented as part of the AMPEL clinical decision
support system.

# Material and methods

## Study population

```{r excluded, include = FALSE}
n_excluded <- attr(raw_data, "excluded")
n_analysed <- nrow(raw_data)
n_total <- sum(n_analysed, n_excluded)

cn_labs <- grep("_[CESQ]$", colnames(raw_data), value = TRUE)
```

In a retrospective cohort study we followed `r n_total` consecutive
patients, who were recruited during the evaluation process for
liver transplantation at the University Hospital of Leipzig from
`r paste(format(attr(raw_data, "period"), "%B %Y"), collapse = " to ")`.
For each patient we recorded `r ncol(raw_data)` variables.
Among them are age, sex, etiology of liver disease, complications as listed in
Table \@ref(tab:tbl1) and `r length(cn_labs)` laboratory measurements.

```{r flowchart, results = "asis", eval = knitr::is_html_output(), echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "Flowchart Inclusion/Exclusion."}
graph <- "
digraph flowchart {
    graph [splines = false];

    // invisible nodes
    node [shape = point, width = 0, height = 0];
    iAR [style = 'invis'];

    // visible nodes
    node [shape = box];
    A [label = '@@1'];
    E [label = '@@3'];
    R [label = '@@2'];

    // edge statements
    A -> iAR [arrowhead = none];
    {rank = same; iAR -> E [minlen = 5]};
    iAR -> R;
}

[1]: paste0('Total Records\\n(n = ', n_total, ')')
[2]: paste0('Analysed Records\\n(n = ', n_analysed, ')')
[3]: str_excluded
"
str_excluded <- paste0(
    'Excluded (n = ', sum(n_excluded), ')\\\\l',
    paste0(
        '- ',
        paste(
            c(
                'Younger than 18 years:',
                'Liver Transplantation before visit:',
                'Follow-up missing/invalid:'
            ),
            n_excluded
        ),
        collapse = '\\\\l'
    ),
    '\\\\l'
)

DiagrammeR::grViz(graph)
```

We excluded `r sum(n_excluded)` patients from our analysis
(Figure \@ref(fig:flowchart)).
`r if (n_excluded["Children"] == 1) "One was" else paste(n_excluded["Children"], "were")` younger than 18 years.
`r n_excluded["LiverTransplantation"]` had a liver transplantation before and
in `r n_excluded["LostFollowUp"]` cases the follow up data were missing or
invalid.

The Ethics Committee at the Leipzig University Faculty of
Medicine approved the retrospective usage of the data for our study
(reference number: 039/14ff).

## MELD scores

MELD and MELD-Na were calculated as described in
@optn_policies2021 using the following formulas:
$MELD = 10 * (0.957 \ln(creatinine [mg/dl]) + 0.378 \ln(bilirubin [mg/dl]) + 1.120 * \ln(INR) + 0.643)$,
Creatinine, bilirubin and INR values lower than 1.0 mg/dl were set to 1.0 mg/dl.
The maximum allowed creatinine was 4.0 mg/dl. If patients received dialysis,
creatinine was set to 4.0 mg/dl.
For MELD-Na MELD was calculated as above and recalculated if greater than 11 using:
$MELD\text{-}Na = MELD_i + 1.32 * (137 - Na [mmol/l]) - (0.033 * MELD_i * (137 - Na [mmol/l]))$

Serum sodium values lower than 125 mmol/l or higher than 137 mmol/l were set
to 125 mmol/l and 137 mmol/l, respectively.

The MELD-Plus7 risk score was calculated as described in @kartoun2017:
$$
\begin{aligned}
L = 8.53499496 &+ \\
    2.59679650 &* \log_{10}(1 + creatinine [mg/dl]) + \\
    2.06503238 &* \log_{10}(1 + bilirubin [mg/dl]) + \\
    2.99724802 &* \log_{10}(1 + INR) - \\
    6.47834101 &* \log_{10}(1 + sodium [mmol/l]) - \\
    6.34990436 &* \log_{10}(1 + albumin [g/l]) + \\
    1.92811726 &* \log_{10}(1 + wbc [th/cumm]) + \\
    0.04070442 &* age [years]
\end{aligned}
$$

$$MELD\text{-}Plus = \frac{\exp(L)}{1 + \exp(L)}$$

The MELD-Plus9 score was ignored because the length of stay information was not
available for our cohort and the model itself wasn't superior to MELD-Plus7 in
its original publication [@kartoun2017].

## Data processing and machine learning

```{r ml, include = FALSE}
n_lrns <- sum(!grepl("^scale..*", bmrk_results$learners$learner_id))
```

All data processing, statistical and machine learning analyses were performed
using R version
`r paste0(sessionInfo()$R.version[c("major", "minor")], collapse = ".")`
[@R-base].
Prior to the analyses all laboratory values were *zlog* transformed to
approximate a normal distribution as described in
@hoffmann2017 and implemented in @R-zlog.

$$
z = (\log(x) - \frac{\log(LL) + \log(UL)}{2}) * \frac{3.92}{\log(UL) - \log(LL)}
$$

Where $LL$ and $UL$ are the lower and upper reference limit of
the corresponding laboratory diagnostic, respectively.
Missing laboratory measurements were replaced with a *zlog* value of zero.
In similar manner if information about complications or
comorbidities were missing we treat them as not present.

We compared `r n_lrns` different statistical and machine learning
algorithms using the *mlr3* framework
[@mlr3; @R-mlr3; @mlr3proba2021; @R-mlr3proba].
The used algorithms are designed for analysis of survival data:
Cox proportional hazards regression model
[@cox1972; @survival-book; @R-survival],
penalized regressions,
namely two different implementations of the
lasso (least absolute shrinkage and selection operator) regression
[@tibshirani1997; @glmnet2011; @R-glmnet; @penalized2010; @R-penalized],
two different implementations of the
ridge regression [@glmnet2011; @R-glmnet; @penalized2010; @R-penalized], and
the elastic net regression [@glmnet2011; @R-glmnet],
two different random forest implementations
[@ranger2017; @R-ranger; @randomForestSRC2008; @R-randomForestSRC],
extreme gradient boosting [@chen2016; @R-xgboost],
two different support vector machines [@vanbelle2011; @R-survivalsvm]
and two neural networks [@katzman2018; @kvamme2019; @R-survivalmodels].

We used nested resampling to avoid feature- and
model-selection-bias to get reliable metrics to choose the best model.
Therefore, we applied a `r crossval$inner$param_set$values$folds`-fold inner
and a `r crossval$outer$param_set$values$folds`-fold outer cross-validation
that was repeated `r crossval$outer$param_set$values$repeats` times.
We searched the hyperparameter space using a grid search with a resolution of
`r tuner$param_set$get_values()$resolution`.
The models were evaluated using the concordance index, which is a measure of
rank correlation between predicted and observed events [@harrell1982].
Subsequently we selected the model with the highest median concordance index.
If multiple models were comparable we chose the simpler (more penalized) one.

After selecting the final model we trained and validated it as described in
@harrell1996.
Briefly we developed our model using a `r ameldcfg$nfold` cross-validation
that was repeated `r ameldcfg$nrepcv` times and finally predicted survival
at `r ameldcfg$times` days.
The predictions were sorted and grouped into equal sized intervals so that there
are `r ameldcfg$m` patients in each interval.
For each group the mean survival probability and the difference to the
Kaplan-Meier estimate was calculated.
The same procedure was repeated for `r ameldcfg$nboot` bootstrap samples.
Subsequently the differences between predicted survival and Kaplan-Meier
estimates of the bootstrap samples were averaged and added to the original
model difference to obtain a bias-corrected estimated between predicted and
observed survival.

The final model was compared against the MELD, MELD-Na and MELD-Plus7 score
using the area under the time-dependent receiver operating characteristic curve
(AUROC) based on the nonparametric inverse probability of censoring weighting
estimator (IPCW) [@blanche2013, @R-timeROC].
Testing was applied as described in @blanche2013.
A $p < 0.05$ was considered as statistical significant difference.

Summary tables were built using the *gtsummary* package
[@gtsummary; @R-gtsummary].

# Results

## Baseline characteristics

```{r tbl1, results = "asis", echo = FALSE, message = FALSE, warning = FALSE}
library("gtsummary")
theme_gtsummary_journal(journal = "jama")
theme_gtsummary_compact()
tbl1
```

```{r tbl1_helper_func, echo = FALSE}
.tbl1 <- function(variable, column = "Overall", ...) {
    id <- gtsummary::inline_text(
        tbl1, variable = all_of(variable), column = "tbl_id1"
    )
    if (column == "Overall")
        column <- "stat_0"
    gtsummary::inline_text(
        tbl1$tbls[[id]],
        variable = all_of(variable), column = all_of(column), ...
    )
}
```

In this study we analysed a cohort of `r n_analysed` patients evaluated for
liver transplantation (Tab. \@ref(tab:tbl1)).
`r sprintf("%i (%s)", sum(raw_data$Sex == "male"), gtsummary::style_percent(mean(raw_data$Sex == "male"), symbol = TRUE))`
of them were men.
The median (IQR) follow-up times was `r .tbl1("DaysAtRisk")` days.
Within the follow-up time `r .tbl1("LTx")` patients received a liver transplant
and were censored for the analysis beyond this time point.
The median (IQR) age of the cohort was `r .tbl1("Age")` years.

Most of the patients presented with cirrhosis
(`r .tbl1("Cirrhosis")`).
The remaining patients had an acute liver failure
(`r .tbl1("ALF")`) or
a chronic liver insufficiency
(`r .tbl1("CLI")`).
At evaluation time the median (IQR) MELD and MELD-Na score were
`r .tbl1("ScoreMeldUnos")` and
`r .tbl1("ScoreMeldNaUnos")`, respectively.
Within 90 days `r .tbl1("M90")` patients died.
In nearly two third of all patients alcohol was the main cause
(`r .tbl1("Ethyltoxic")`) of their end-stage liver disease.
The remaining patients had viral infections
(hepatitis B and C viruse `r .tbl1("HBV")` and `r .tbl1("HCV")`, respectively),
autoimmune processes (autoimmune hepatitis `r .tbl1("AIH")`,
primary sclerosing cirrhosis `r .tbl1("PSC")`),
primary biliary cirrhosis `r .tbl1("PBC")` or unknown causes
(non-alcoholic steatohepatitis `r .tbl1("NASH")` and
cryptogenic hepatitis `r .tbl1("Cryptogenic")`).
The most common complications among all patients were gastrointestinal
bleedings in `r .tbl1("GIB")` patients, followed by
hepatocellular carcinoma (HCC) in `r .tbl1("HCC")` and
spontaneous bacterial peritonitis (SBP) in `r .tbl1("SBP")` patients.
Just `r .tbl1("Dialysis")` patients required a renal dialysis.

## MELD scores

According to @wiesner2003 we divided our cohort into
`r nlevels(labelled_meld_data$MeldCategory)` MELD-score risk
categories, with `r enum(table(labelled_meld_data$MeldCategory))` patients,
respectively.
As shown in Table \@ref(tab:tbl-observed-vs-expected-mortality)
our observed 90-day mortality was
`r paste(round(range(tbl_observed_vs_expected_mortality$StandardizedMortalityRatio[-1]), 1), collapse = " to ")`
times higher than the predicted one. Just in the group with the lowest MELD
scores the mortality was much lower.

```{r tbl-observed-vs-expected-mortality, results = "asis", echo = FALSE, message = FALSE}
tbl_observed_vs_expected_mortality$ObservedMortality <-
    tbl_observed_vs_expected_mortality$ObservedMortality * 100
tbl_observed_vs_expected_mortality$ExpectedMortality <-
    tbl_observed_vs_expected_mortality$ExpectedMortality * 100
knitr::kable(
    cbind(
        row.names(tbl_observed_vs_expected_mortality),
        tbl_observed_vs_expected_mortality
    ),
    row.names = FALSE,
    col.names =
        c(
            "MELD category",
            "Observed deaths (n)",
            "Expected deaths (n)",
            "Standardized mortality ratio (SMR)",
            "Observed mortality (%)",
            "Expected mortality (%)"
        ),
    caption = paste0(
        "Observed vs MELD-expected 90-day mortality. ",
        "MELD mortality values are taken from @wiesner2003. ",
        "All patients censored before day 90 are ignored for ",
        " the calculation of the MELD-expected deaths. ",
        "SMR, Standardized mortality ratio = observed deaths/expected deaths."
    ),
    digits = 1
)
```

## AMPEL MELD

The penalized regression and random forest models
result in a similar high median performance in the benchmark process.
In favour of simpler models we chose a lasso based approach for the final model
development.
The classical cox model, the data and computational intense algorithms
extreme gradient boosting, support vector machines and neural networks yield a
worse performance.

```{r coeftbl, echo = FALSE}
tar_load(bootrcv)
cf <- coef(bootrcv$fit)
cf <- setNames(cf@x, rownames(cf)[cf@i + 1])
o <- order(-abs(cf))
cfs <- cf[o]
cfnms <- lapply(
    labelled_meld_data[names(cfs)],
    function(nm)sub(" *\\[.*$", "", attr(nm, "label"))
)
cfnms[!lengths(cfnms)] <- names(cfnms)[!lengths(cfnms)]
knitr::kable(
    setNames(cfs, cfnms),
    col.names = "Coefficient",
    caption = "Model coefficients",
    digits = 3
)
```

The resulting model AMPEL model of end-stage liver disease
(AMELD) uses `r length(cfs)` predictors.
Beside the classical MELD variables INR and (direct) bilirubin the lasso
regression selected cystatin C over creatinine, and
additionally IL-6, total protein and cholinesterase.

```{r basehaz, echo = FALSE}
bh <- basehaz(
    bootrcv$fit, x = amelddata$x, y = amelddata$y, s = "lambda.1se",
    times = timeROC_RCV$times, centered = FALSE
)
s0 <- setNames(1 - bh$hazard, bh$time)
k <- knitr::kable(
    t(s0),
    digits = 4,
    caption = paste0(
        "Underlying survival function; S~0~: baseline survival estimate"
    )
)
k[3] <- paste("|t [days]", k[3])
k[4] <- paste0("|:------", k[4])
k[5] <- paste("|S~0~ (t)", k[5])
k
```

The 90-day survival estimate of a given patient could be calculated as shown in
\@ref(eq:survcalc).

```{r survivalcalc, results = "asis", echo = FALSE}
cfnmsltx <- paste0("\\text{", gsub(" ", "\\ ", cfnms), "}")
cat(
    "$$\\begin{align}",
    "S &= S_{0,90}^{\\exp{LP}} \\\\",
    "&= ", round(s0["90"], 4), "^{\\exp{(",
        paste0(
            round(cfs[1], 4), " * zlog(", cfnmsltx[1], ")",
            paste0(
                ifelse(sign(cfs[-1]) > 0, "+", ""),
                round(cfs[-1], 4), " * zlog(", cfnmsltx[-1], ")",
                collapse = ""
            ),
            collapse = ""
        ),
    ")}}"
)
```
(\#eq:survcalc)
\end{align}
$$
<!-- using eq:survcalc inside the R code chunk impedes Math ML rendering -->

Where $S_{0}$ is the baseline survival estimate from
\@ref(tab:basehaz) and $LP$ is the linear predictor of the coefficients
in \@ref(tab:coeftbl).

## Model comparison

```{r roc, fig.width = 7, fig.height = 7, echo = FALSE, message = FALSE}
col <- viridisLite::viridis(5)
m <- list(
    AMELD = timeROC_RCV,
    MELD = timeROC_MELD,
    "MELD-Na" = timeROC_MELDNa,
    "MELD-Plus7" = timeROC_MELDPlus7
)
invisible(roc <- plot_surv_roc(m, timepoint = 90, col = col[-2]))
roc <- roc[order(-roc)]
```

```{r roctest, echo = FALSE}
tp <- timeROC::compare(
    timeROC_RCV, timeROC_MELDNa,
    adjusted = TRUE
)$p_values_AUC
rownames(tp) <- paste(
    c("Non-adjusted", "Adjusted"), "$p$-values"
)
knitr::kable(
    tp,
    caption = paste0(
        "Test comparison between AUCROC for MELD-Na and AMELD ",
        "according to @blanche2013 at day ",
        paste0(timeROC_RCV$times, collapse = ", "), ".",
        "$p$-values are shown as non-adjusted and adjusted for ",
        "multiple testing."
    ),
    digits = 2
)
```

AMELD offers the best AUROC (`r round(roc["AMELD"], 3)`) across all competitors.
However, after correction for multiple testing the difference against
the second best model `r names(roc)[2]`
(AUROC: `r round(roc[2], 3)`) is not statistical significant
(`r paste0("$p = ", round(tp[2, "t=90"], 2), "$")`)
(see Table \@ref(tab:roctest) for complete comparison).

```{r bootrcv, echo = FALSE, fig.width = 8, fig.height = 6, fig.cap = paste('Calibration of AMELD in comparison to MELD, MELD-Na and MELD-Plus7.\n The points mark the mean predicted survival of', ameldcfg$m, 'patients per interval against their observed mean survival. The error bars correspond to the 95 % confidence interval of the survival estimate.'), echo = FALSE}
par(cex = 0.9)
plot(
    bootrcv,
    what = "calibration",
    main = "90-day survival",
    legend = FALSE
)

ps <- lapply(
    zlog_data[paste0("SurvProbMeld", c("Unos", "NaUnos", "Plus7"))],
    function(p) {
        ctpnts <- cutpoints(p, n = ameldcfg$m)
        f <- cut(p, ctpnts, include.lowest = TRUE)
        list(
            predicted = groupmean(p, f = f),
            observed = observed_survival(
                amelddata$y, f = f, times = ameldcfg$times
            )
        )
    }
)
names(ps) <- c("MELD", "MELD-Na", "MELD-Plus7")

for (i in seq_along(ps)) {
    lines(
        ps[[i]]$predicted, ps[[i]]$observed, col = col[i + 2],
        type = "b", pch = 19
    )
}
legend(
    "topleft",
    col = col,
    legend = c(
        "AMELD",
        paste0(
            "AMELD, optimism error added (based on ",
            length(bootrcv$models),
            " bootstrap samples)"
        ),
        names(ps)
    ),
    pch = c(19, 4, rep(19, 3)),
    bty = "n"
)
```

Fig. \@ref(fig:bootrcv) depicts the calibration curve for AMELD and the
competing MELD scores.
While the prediction for low-risk patients is similar to other scores
AMELD predicts the medium-risk patients more accurate.
Because of the low number of very ill patients in our cohort the prediction
performance in high-risk patients is poor but comparable to the other models.
This would require a recalibration in a larger validation cohort.

# Discussion

In this retrospective analysis of `r n_analysed` consecutive patients,
who were recruited during the evaluation process for liver transplantation at
the University Hospital of Leipzig we developed a new model to predict the
90-day survival probability. Our new model, AMELD, outperforms the MELD-Na
score and the recently published MELD-Plus7 risk score.

```{r selvar, echo = FALSE}
bsrank <- as.vector(rank(-ameld:::.selvars(bootrcv$models, bootrcv$s), ties = "first")["NA_S"])

bsranktxt <-
    if (!is.na(bsrank)) {
        paste0("in bootstrap selection rank ", bsrank, " of ", ncol(amelddata$x))
    } else {
        paste0("it was never selected in any boostrap sample")
    }

rfrank <- as.vector(rank(-rngr$variable.importance, ties = "first")["NA_S"])
rfranktxt <-
    paste0("in a random forest variable importance rank ", rfrank, " of ", ncol(amelddata$x))
```
Although the Organ Procurement and Transplantation Network recommends the use
of MELD-Na over MELD the sodium does not add any information to the score.
Neither in the bootstrap selection nor in the variable importance of a
random forest model the sodium gains any important weight
(`r bsranktxt` and `r rfranktxt`, see \@ref(fig:vimpbs), \@ref(fig:vimprf)).
In addition sodium levels can be affected by diuretic therapy and already
@kartoun2017 have shown that the discrimination performance of MELD-Na is
insignificantly different from the classical MELD.
Consequentially it is not a component of our model.

It has shown that women receive in median lower scores in MELD-Na and therefore
are less likely getting an urgent listing for liver transplantation
[@gaertner2022].
To address sexual differences in laboratory measurements we applied the *zlog*
transformation [@hoffmann2017] in our study.
Additionally this transformation yield comparable laboratory measurements
across different clinical laboratories and methods.
It also reduce the influence of age- and sex-related differences,
which could improve the objectivity of the allocation process.
The already nearly normal scaled *zlog*-transformed laboratory measurements
allow us to avoid the sample depending scaling prior to model development.
Both should increase the external applicability of our approach.

Our model, AMELD, uses `r length(cfs)` predictors.
Half of them, INR, (direct) bilirubin, and cystatin C are identical
or similar to the labortory measurements utilized in the classical MELD.
The direct bilirubin selected by our approach and the total bilirubin in MELD
are highly correlated (spearman correlation:
`r round(cor(amelddata$x, method = "spearman")["BILI_S", "BILID_S"], 2)`) and
therefore interchangeable.

In our feature selection process cystatin C was chosen instead of creatinine.
This is inline with recent studies that show a better prediction of mortality
for cystatin C in unselected and critical ill patients
[@helmerssonkarlqvist2016; @helmerssonkarlqvist2021]

Furthermore AMELD extends MELD by additional synthesis parameters
total protein and cholinesterase and
most importantly the inflammatory marker IL-6.

The latter is an example where results from machine learning can aid in
finding clinical meaningful patterns and guide further research.
The systemic inflammatory response syndrome is an
independent risk factor for poor outcome in patients with cirrhosis
[@thabut2007].
However neither MELD nor MELD-Na taking inflammation into account.
In contrast IL-6 is the most important variable in our model. Even if IL-6 is
excluded the algorithms select CRP as one of the most important variables which
highlights the important role of inflammation in the prognosis of patients
with end-stage liver disease.
We and other researchers have already shown that adding inflammatory markers,
namely IL-6 or CRP, to the classical MELD score increase the accuracy of the
90-day mortality prediction [@remmler2018; @cervoni2016].

Beside laboratory measurements we investigated
`r (ncol(raw_data) - length(cn_labs) - 2)`
clinical data but none of them were chosen in our model selection process.
@kartoun2017 also used the lasso approach and neither select any clinical
data. Especially in retrospective analysis the clinical data are often not
detailed enough to keep valuable information.

In our benchmark of `r n_lrns` different machine learning algorithms we have
seen that the more complex algorithm like boosting, support vector machines
and neural networks perform worse than simpler ones.
This may due additive effects that favor regression models or due to our
relative small sample size.
It has been shown that in machine learning more than 200 events per
predictor variable are required to achieve good prediction performance
[@ploeg2014].

<!--
As we shown above machine learning can aid in finding clinical meaningful
patterns.
-->

<!--

- higher mortality than predicted, compare @hernaez2020
- adding sodium doesn't add much/any discrimination information
    - as already seen in @kartoun2017
- @kartoun2017 uses > 60 variables mostly clinical stuff, just few laboratory values, none of the clinical stuff remains
- general low sample size thus simpler models/fewer predictors are preferred.
- comparison of different predictors of MELD-Plus7 (+MELD-Plus9) vs our features
    - 90 day mortality *post* discharge
    - poorer performance in own validation set (lower and more uniform disease severity, also lower MELD)
- machine learning algorithms worse than statistical learning (lower c-index, higher variance/instability)
    - additive effects (favour statistical models)
    - low sample size, for ML more than 200 events per predictor variable necessary [@ploeg2014]
- sex-differences and sample-dependencies (by scaling) addressed by *zlog*.
- more variables, harder to cheat
- national register required
-->

## Limitations

Our study has several limitations.
Principally it is a single center, retrospective analysis and without an
external validation cohort.
The data were collected during the evaluation process for liver transplantation
at the University Hospital of Leipzig.
There was no controlled follow-up process.
We excluded all patients who had already a liver transplantation
which could artificial lower the disease severity of the cohort.
Furthermore most of our patients suffered from ethyltoxic liver cirrhosis
(`r .tbl1("Ethyltoxic", pattern = "{p}%")`) which may
limited the transfer to other etiologies.

With a median MELD score of `r .tbl1("ScoreMeldUnos")` our
cohort was healthier than in previous studies that reported
a score of 14-15 [@kim2008; @kartoun2017].
However the 90-day mortality rate of `r .tbl1("M90")` was comparable to
the mortality rates seen in earlier studies that ranges from 6 to 30%
[@malinchoc2000; @kim2008; @kartoun2017].
Although our cohort was healthier with a medium mortality rate the classical
MELD score underestimates the observed mortality
`r paste(round(range(tbl_observed_vs_expected_mortality$StandardizedMortalityRatio[-1]), 1), collapse = " to ")`
times.
Unfortunately our sample size is low in the high risk patient category
which yield a poor but better than MELD based prediction performance of our
AMELD for very ill patients.

## Conclusion

In conclusion we developed a new model to predict the 90-day survival of
patients during the evaluation process for liver transplantation.
Our model, AMELD, extends the classical MELD predictors, INR,
(direct) bilirubin, cystatin C (instead of creatinine) by
synthesis parameters total protein and cholinesterase and
most importantly the inflammatory marker IL-6.
Using these `r length(cfs)` laboratory measurments the AMELD
outperforms the established MELD-Na and the recently published
MELD-Plus7 risk scores.
For a wider adoption a prospective multi center study for model
recalibration and validation is needed.

<!--
Mortality
[@malinchoc2000] ~ 30 %
[@kim2008] ~ 6-7 %
[@kartoun2017] ~ 16.3 % (90 d, post-discharge-mortality)

random time point for blood sampling

It was not possible to ensure that patients died of end-stage liver disease.

The variables we found and the model itself have to be validated and
recalibrated in a multi center, prospective trial.

Few patients with high MELD scores, poor calibration for high mortality

- comparable mortality?
- retrospective analysis
- single center
- unknown cause of death (death with vs of end-stage liver disease)
- bias caused by LTx censorship?
-->

# Data and software availability

The complete code and reproducible analysis could be found at
https://ampel-leipzig.github.io/ampel-leipzig-meld/.
Additionally all data are available in the `ameld` R package [@R-ameld] at
https://github.com/ampel-leipzig/ameld
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.6666018.svg)](https://doi.org/10.5281/zenodo.6666018)
for further investigations.

# Acknowledgement

The study was supported by the AMPEL project ([www.ampel.care](https://www.ampel.care)).
This project is co-financed through public funds according to the budget
decided by the Saxon State Parliament under the
RL eHealthSax 2017/18 grant number 100331796.
The funder provided support in the form of
salaries for author SG
but did not have any additional role in the study design,
data collection and analysis, decision to publish, or preparation
of the manuscript.
We thank the whole AMPEL team for their support,
Stefan Kemnitz for his technical support regarding the use of the
HPC cluster of the University Greifswald and
Bernd Klaus for helpful discussions about survival analysis and proof reading the manuscript.

```{r co2e, include = FALSE}
tdp <- c(
    ## brain/(batch|login) Intel(R) Xeon(R) CPU E5-2650 v4 @ 2.20GHz
    ## https://ark.intel.com/content/www/us/en/ark/products/91767/intel-xeon-processor-e52650-v4-30m-cache-2-20-ghz.html
    login = 105,
    ## brain/snowball Intel(R) Xeon(R) Gold 6240 CPU @ 2.60GHz
    ## https://www.intel.com/content/www/us/en/products/sku/192443/intel-xeon-gold-6240-processor-24-75m-cache-2-60-ghz.html
    ## TDP 150, contains 2 processors
    ## (but even with workers = 72, we barely use more than 36, so 150 instead of 300 W should be fine)
    snowball = 150
)

## Umweltbundesamt
## Climate Change 45/2021 Entwicklung der spezifischen Kohlendioxid-Emissionen des deutschen Strommix in den Jahren 1990 - 2020
## Erscheinungsjahr Mai 2021, Autor(en): Petra Icha, Dr. Thomas Lauf, Gunter Kuhs
## p. 8
## https://www.umweltbundesamt.de/sites/default/files/medien/5750/publikationen/2021-05-26_cc-45-2021_strommix_2021_0.pdf
kgco2kwh <- 0.471

tm <- tar_meta(fields = c("name", "type", "seconds"))
tm <- tm[tm$type %in% c("pattern", "stem"),]
tm$node <-
    ifelse(grepl("^boot|^bmrks$|^arcvob$", tm$name), "snowball", "login")

tm$runtime <- tm$seconds *
    ifelse(tm$node == "snowball", 32, 1) # on snowball we use 32 workers
kgco2e <- sum((tm$runtime / 3600) * (tdp[tm$node] / 1000)) * kgco2kwh
kgco2e

## Umweltbundesamt
## Ökologische Bewertung von Verkehrsarten
## Erscheinungsjahr August 2020, Autor(en): Michel Allekotte, Fabian Bergk, Kirsten Biemann, Carolin Deregowski, Wolfram Knörr, Hans-Jörg-Althaus, Daniel Sutter, Thomas Bergmann
## p. 214-215
## https://www.umweltbundesamt.de/sites/default/files/medien/479/publikationen/texte_156-2020_oekologische_bewertung_von_verkehrsarten_0.pdf
## Gesamt CO2eq (nur Verbrauch und Energiebereitstellung) je Personen km,
## A2, Tab. 69: PKW TTW 131 gCO2eq/Pkm, A3, Tab. 71.: PKW WTT 22 gCO2eq/Pkm
carkgco2e <- 153 / 1000
km <- kgco2e / carkgco2e
```

The calculations were conducted using the high-performance computing (HPC)
cluster of the University Greifswald.
In 2018 the average carbon efficiency of the German power grid was
`r kgco2kwh` kgCO~2~eq/kWh [@icha2021].
A single run of the complete calculation pipeline took roughly
`r round(sum(tm$runtime) / 3600)` hours of computation on an
Intel(R) Xeon(R) E5-2650 v4 and
Intel(R) Xeon(R) Gold 6240 CPU (thermal design power: 105 and 150 W).
The total estimated emissions were at least
`r sprintf("%.2f", kgco2e)` kgCO~2~eq
(ignoring development, test runs, etc.).
This is equivalent to `r round(km)` km driven by an average car
[@allekotte2021].

# References